---
title: "Ed's frog occupancy and abundance analysis"
author: "Ed Basham & Alex Baecher"
date: "2/19/2021"
output: html_document
---
# Ed's frog analysis
## Alex's help with some data / modeling stuff
```{r Preparing libraries, message=FALSE, warning=FALSE}
## Prepare packages

pacman::p_load(
  "ade4",
  "matrixStats",
  "fields",
  "Hmisc",
  "MuMIn",
  "ggpubr",
  "TMB",
  "glmmTMB",
  "visreg",
  "devtools",
  "tidyverse",
  "lme4",
  "AICcmodavg",
  "performance",
  "see",
  "insight")

```

## Load data
```{r Loading data, message=FALSE, warning=FALSE}
### Frog data
frog_dat <- read_csv("frog_dat.csv") # this is the frog data, which contains all the information from frog captures during surveys

### Survey data
survey_dat <- read_csv("survey_dat.csv") # this is the survey data, which contains all information from frog surveys

### Tree characteristics data
tree_dat <- read_csv("tree_dat.csv") # this is the survey data, which contains all information about trees the surveys were conducted on
```

## Merging together all three dataframes
```{r Merging three dataframes}
### Combining survey and tree data
survey_tree_dat <- survey_dat %>%
  left_join(tree_dat) 

### Summarizing frog data by survey
frog_dat_indnum_sum <- frog_dat %>%
  group_by(tree_code, date, am_pm) %>%
  summarize(count = length(unique(ANFU_code)))

### Combining frog data with previously combined tree-survey data 
global_dat <- survey_tree_dat %>%
  left_join(frog_dat_indnum_sum) %>%
  mutate(count = replace_na(count, 0),
         pres = if_else(count > 0, 1, count),
         geography = scale(latitude + longitude),
         espave_labels = factor(espave, 
                                levels = c(0,1),
                                labels = c("Non-Espave", "Espave")),
         arboreal_survey = ifelse(str_detect(survey_types, "A"), "Arboreal", NA),
         ground_survey = ifelse(str_detect(survey_types, "G"), "Ground", NA),
         day = scale(as.numeric(lubridate::dmy(global_dat$date) - lubridate::dmy("27/06/2017")))) %>%
  relocate(tree_code, Season, date, am_pm, pres, count, arboreal_survey, ground_survey); global_dat

```

## Creating supplimental table
```{r}
tree_supp <- global_dat %>% 
  group_by(tree_species,
           Season) %>%
  summarize(arboreal_survey_count = sum(!is.na(arboreal_survey)),
            ground_survey_count = sum(!is.na(ground_survey))) %>%
  pivot_wider(id_cols = c(tree_species,Season,arboreal_survey_count,ground_survey_count),
              names_from = Season, 
              values_from = c(arboreal_survey_count,ground_survey_count),
              values_fill = 0) %>%
  rename(`Dry Arboreal` = arboreal_survey_count_D,
         `Wet Arboreal` = arboreal_survey_count_W,
         `Dry Ground` = ground_survey_count_D,
         `Wet Ground` = ground_survey_count_W) %>%
  left_join(global_dat %>% 
              group_by(tree_species,
                       tree_common_name) %>%
              summarize(tree_count = n_distinct(tree_code)))

```

## Calculating nearest neighbor distance to occupied trees
```{r}
## Create function for calculating distance using spatial geometry
earth_dist<-function(lat1,long1,lat2,long2){

  rad <- pi/180
  a1 <- lat1 * rad
  a2 <- long1 * rad
  b1 <- lat2 * rad
  b2 <- long2 * rad
  dlat <- b1-a1
  dlon<- b2-a2
  a <- (sin(dlat/2))^2 +cos(a1)*cos(b1)*(sin(dlon/2))^2
  c <- 2*atan2(sqrt(a),sqrt(1-a))
  R <- 6378.145
  dist <- R *c
  return(dist)
}

### Filtering data to occupied trees
global_pres_dat <- global_dat %>%
  filter(pres == 1)

### Initializing matrix
dist <- matrix(NA, ncol = nrow(global_pres_dat), nrow = nrow(global_dat))

### Looping over matrix to calculate nearest neighbor using function
for (i in 1:nrow(global_dat)){
  for(j in 1:nrow(global_pres_dat)){
    dist[i,j] <- earth_dist(global_dat$latitude[i],
                            global_dat$longitude[i],
                            global_pres_dat$latitude[j],
                            global_pres_dat$longitude[j])
  }}

{
  d <- as.data.frame(dist*1000);
  d <- d[,1:nrow(global_pres_dat)]
}

nn.dist <- apply(d, 1, function(x) {
  return(sort(x, partial = 2)[2])
})

global_dat$nn_dist <- scale(nn.dist)

```


# Prepare data for presence and counts modeling
```{r Presence models}
# pres_pseudor2 <- r.squaredGLMM(glmer(pres ~ 1 + (1|tree_code), family = binomial, data = global_dat))           
# pres_pseudor2_geo <- r.squaredGLMM(glmer(pres ~ 1 + geography + (1|tree_code), family = binomial, data = global_dat))  
# pres_pseudor2 # delta: cond = 0.5972025, marg = 0
# pres_pseudor2_geo # delta: cond = 0.6920702, marg = 0.4630165 
# 
# 0.6920702 - 0.5972025 # difference in delta; 0.4007134
```


# Create models for presence absence
```{r}
pres_mods_formulas <-  c(
  "pres ~ 1 +                          geography + day_hour + (1|tree_code)", 
  "pres ~ epiphytes +                  geography + day_hour + (1|tree_code)", 
  "pres ~ canopy_cover +               geography + day_hour + (1|tree_code)", 
  "pres ~ ground_cover +               geography + day_hour + (1|tree_code)",
  "pres ~ canopy_height +              geography + day_hour + (1|tree_code)",
  "pres ~ large_vines +                geography + day_hour + (1|tree_code)",
  "pres ~ moss +                       geography + day_hour + (1|tree_code)",
  "pres ~ dbh +                        geography + day_hour + (1|tree_code)",
  "pres ~ root_base +                  geography + day_hour + (1|tree_code)",
  "pres ~ lugovia_epiphytes +          geography + day_hour + (1|tree_code)",
  "pres ~ stream_distance +            geography + day_hour + (1|tree_code)",
  "pres ~ density_small_stems +        geography + day_hour + (1|tree_code)",
  "pres ~ density_big_stems +          geography + day_hour + (1|tree_code)",
  "pres ~ espave +                     geography + day_hour + (1|tree_code)",
  "pres ~ canopy_cover*espave +        geography + day_hour + (1|tree_code)",
  "pres ~ ground_cover*espave +        geography + day_hour + (1|tree_code)",
  "pres ~ canopy_height*espave +       geography + day_hour + (1|tree_code)", #prob
  "pres ~ large_vines*espave +         geography + day_hour + (1|tree_code)",
  "pres ~ moss*espave +                geography + day_hour + (1|tree_code)",
  "pres ~ dbh*espave +                 geography + day_hour + (1|tree_code)",
  "pres ~ epiphytes*espave +           geography + day_hour + (1|tree_code)", #prob
  "pres ~ root_base*espave +           geography + day_hour + (1|tree_code)",#prob
  "pres ~ lugovia_epiphytes*espave +   geography + day_hour + (1|tree_code)",#prob
  "pres ~ stream_distance*espave +     geography + day_hour + (1|tree_code)")#prob

pres_mods <- lapply(pres_mods_formulas, function(x){
  glmer(as.formula(x), family = binomial, data = global_dat)})

pres_mods_output <- aictab(pres_mods, pres_mods_formulas) %>% 
  as.data.frame()

pres_R2_nakagawa <- data.frame(
  cond_R2 = lapply(pres_mods, function(x){
    r2_nakagawa(x)[1]}) %>%
    unlist(),
  marg_R2 = lapply(pres_mods, function(x){
    r2_nakagawa(x)[2]}) %>%
    unlist())
pres_R2_nakagawa

pres_mods_output$marg_R2 <- pres_R2_nakagawa$marg_R2
pres_mods_output$cond_R2 <- pres_R2_nakagawa$cond_R2

pres_mods_output %>%
  select(Modnames, K, Delta_AICc, AICcWt, Cum.Wt, marg_R2, cond_R2) 

pres_performance <- compare_performance(pres_mods, rank = T) %>%
  data.frame() %>%
  mutate(Model = pres_mods_output$Modnames)

```

# Checking performance of top model
```{r}
top_pres_mod <- glmer(as.formula(pres_mods_output$Modnames[1]), family = binomial, data = global_dat)
  
r2_nakagawa(top_pres_mod, by_group = T)

check_model(top_pres_mod)

model_performance(top_pres_mod)

check_overdispersion(top_pres_mod)

check_singularity(top_pres_mod)
```


## Plotting model predictions
```{r Plotting presence model predictions}
global_dat_pres_dbh <- cbind(preds = glmer(as.formula(pres_mods_output$Modnames[1]), family = binomial, data = global_dat) %>% 
  predict(type = "response"), global_dat)
global_dat_pres_moss_espave <- cbind(preds = glmer(as.formula(pres_mods_output$Modnames[2]), family = binomial, data = global_dat) %>% 
  predict(type = "response"), global_dat)
global_dat_pres_espave <- cbind(preds = glmer(as.formula(pres_mods_output$Modnames[3]), family = binomial, data = global_dat) %>% 
  predict(type = "response"), global_dat)

gg_pres_dbh <- ggplot(data = global_dat_pres_dbh, aes(x = dbh, y = preds, col = "All")) + 
  stat_smooth(size = 1, alpha = 0.5, method = "glm", method.args = list(family = "quasibinomial")) +
  xlab("Diameter at Breast Height (scaled)") + ylab("Probability of Presence") +
  scale_color_manual("Tree Species", values = c("firebrick3"), labels = c("All")) + 
  coord_cartesian(ylim = c(0,1)) + theme_classic() + theme(legend.position = "none"); gg_pres_dbh

gg_pres_moss_espave <- ggplot(global_dat_pres_moss_espave, aes(x = espave, y = preds, col = espave, fill = espave)) + 
  geom_boxplot(alpha = 0.7, width = 0.5, size = 0.75) + xlab(" ") + ylab("Probability of Presence") + 
  scale_x_discrete(labels = c("Non-Espavé","Espavé")) +
  scale_color_manual("Tree Species", values = c("orange","grey20"), labels = c("Non-Espavé","Espavé")) + 
  scale_fill_manual("Tree Species", values = c("orange","grey78"), labels = c("Non-Espavé","Espavé")) + 
  facet_wrap(~factor(moss, levels = c(0, 1), labels = c("No moss", "Moss"))) +
  theme_classic() + theme(legend.position = "none", 
                          strip.text.x = element_text(face = "bold", size = 11),
                          axis.text.x = element_text(size = 10.92, color = "black", vjust = -4)); gg_pres_moss_espave

gg_pres_espave <- ggplot(global_dat_pres_espave, aes(x = espave, y = preds, fill = espave)) + 
  geom_boxplot(alpha = 0.7, width = 0.5, size = 0.75) + xlab(" ") + ylab("Probability of Presence") + 
  scale_x_discrete(labels = c("Non-Espavé","Espavé")) +
  scale_fill_manual("Tree Species", values = c("orange","grey78"), labels = c("Non-Espavé","Espavé")) + 
  theme_classic() + theme(legend.position = "none", axis.text.x = element_text(size = 10.92, color = "black", vjust = -4)); gg_pres_espave

pres_3_pan <- ggarrange(gg_pres_dbh, gg_pres_moss_espave, gg_pres_espave,
          ncol = 1, labels = c("(a)","(b)","(c)"),
          label.x = 0.12, label.y = 0.985); pres_3_pan
```

```{r}
count_prob_dist <- list()
count_prob_dist[[1]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~1, family=poisson, data = global_dat)
# zero-inflated Poisson model with a single zeroinflation parameter applying to all observations

count_prob_dist[[2]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~0, family=poisson, data = global_dat)
# zero-inflated Poisson model with an excluded single zeroinflation parameter

count_prob_dist[[3]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~1, family = nbinom1(link = "log"), data = global_dat)
# zero-inflated negative binomial model; NB parameterization

count_prob_dist[[4]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~0, family = nbinom1(link = "log"), data = global_dat)
# zero-inflated negative binomial model; NB parameterization with an excluded single zeroinflation parameter

count_prob_dist[[5]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~1, family = nbinom2(link = "log"), data = global_dat)
# zero-inflated negative binomial model; the default is the “NB2” parameterization (variance = µ(1 + µ/k): Hardin and Hilbe (2007))

count_prob_dist[[6]]  <- glmmTMB(count ~ 1 + (1|tree_code), ziformula=~0, family = nbinom2(link = "log"), data = global_dat)
# zero-inflated negative binomial model; with an excluded single zeroinflation parameter

count_prob_dist_output <- aictab(cand.set = count_prob_dist, modnames = c("poisson, ziform=~1",
                                                                          "poisson, ziform=~0",
                                                                          "nbinom1, ziform=~1",
                                                                          "nbinom1, ziform=~0",
                                                                          "nbinom2, ziform=~1",
                                                                          "nbinom2, ziform=~0"))

# nbinom1, ziform=~0 was the top ranking model according to AICc, but the R2 values were approx = 1 (problematic). 
# The rbinom2, ziform=~0 model had lower R2 values, so I'm proceeding with it

```

# Create models for counts
```{r}
count_mods_formulas <- c(
  "count ~ 1 +                        geography + day_hour + (1|tree_code)", 
  "count ~ epiphytes +                geography + day_hour + (1|tree_code)", 
  "count ~ mean_canopy_cover +        geography + day_hour + (1|tree_code)", 
  "count ~ mean_ground_cover +        geography + day_hour + (1|tree_code)",
  "count ~ canopy_height +            geography + day_hour + (1|tree_code)",
  "count ~ large_vines +              geography + day_hour + (1|tree_code)",
  "count ~ moss +                     geography + day_hour + (1|tree_code)",
  "count ~ dbh +                      geography + day_hour + (1|tree_code)",
  "count ~ root_base +                geography + day_hour + (1|tree_code)",
  "count ~ lugovia_epiphytes +        geography + day_hour + (1|tree_code)",
  "count ~ stream_distance +          geography + day_hour + (1|tree_code)",
  "count ~ density_small_stems +      geography + day_hour + (1|tree_code)",
  "count ~ density_big_stems +        geography + day_hour + (1|tree_code)",
  "count ~ espave +                   geography + day_hour + (1|tree_code)",
  "count ~ mean_canopy_cover*espave + geography + day_hour + (1|tree_code)",
  "count ~ mean_ground_cover*espave + geography + day_hour + (1|tree_code)",
  "count ~ canopy_height*espave +     geography + day_hour + (1|tree_code)",
  "count ~ large_vines*espave +       geography + day_hour + (1|tree_code)", 
  "count ~ moss*espave +              geography + day_hour + (1|tree_code)",
  "count ~ dbh*espave +               geography + day_hour + (1|tree_code)",
  "count ~ stream_distance*espave +   geography + day_hour + (1|tree_code)")

# fit with default (nlminb) and alternative (optim/BFGS) optimizer
count_mods <- lapply(count_mods_formulas, function(x){
  glmmTMB(as.formula(x), ziformula=~0, family = nbinom2, data = global_dat)})

count_mods_output <- aictab(count_mods, count_mods_formulas) %>% 
  as.data.frame()

count_R2_nakagawa <- data.frame(
  cond_R2 = lapply(count_mods, function(x){
    r2_nakagawa(x)[1]}) %>%
    unlist(),
  marg_R2 = lapply(count_mods, function(x){
    r2_nakagawa(x)[2]}) %>%
    unlist())
count_R2_nakagawa

count_mods_output$marg_R2 <- count_R2_nakagawa$marg_R2
count_mods_output$cond_R2 <- count_R2_nakagawa$cond_R2

count_mods_output

count_mods_output %>%
  select(Modnames, K, Delta_AICc, AICcWt, Cum.Wt, marg_R2, cond_R2) %>%
  write_csv("count_mods_output.csv")

count_performance <- compare_performance(count_mods, rank = T) %>%
  data.frame() %>%
  mutate(Model = count_mods_output$Modnames)

```

# Check performance of top model
```{r}
top_count_mod <- glmer(as.formula(count_mods_output$Modnames[17]), family = binomial, data = global_dat)
  
r2_nakagawa(top_count_mod, by_group = T)

check_model(top_count_mod)

model_performance(top_count_mod)

check_overdispersion(top_count_mod)

check_singularity(top_count_mod)
```

```{r}
global_dat_count_canopy_espave <- cbind(preds = glmmTMB(as.formula(count_mods_output$Modnames[1]), ziformula=~0, family = nbinom2, data = global_dat) %>% 
  predict(type = "response"), global_dat)
global_dat_count_espave <- cbind(preds = glmmTMB(as.formula(count_mods_output$Modnames[2]), ziformula=~0, family = nbinom2, data = global_dat) %>% 
  predict(type = "response"), global_dat)
global_dat_count_dbh <- cbind(preds = glmmTMB(as.formula(count_mods_output$Modnames[3]), ziformula=~0, family = nbinom2, data = global_dat) %>% 
  predict(type = "response"), global_dat)

gg_count_canopy_espave <- ggplot(data = global_dat_count_canopy_espave, aes(x = canopy_height, y = preds, col = espave)) + 
  stat_smooth(size = 1, alpha = 0.5, method = "glm", method.args = list(family = "quasipoisson")) + coord_cartesian(ylim = c(0,5)) +
  xlab("Canopy Height (scaled)") + ylab("Predicted Abundance") + 
  scale_color_manual("Tree Species", values = c("orange","grey25"), labels = c("Non-Espavé","Espavé")) + 
  theme_classic() + theme(legend.position = "none"); gg_count_canopy_espave

gg_count_espave <- ggplot(global_dat_count_espave, aes(x = espave, y = preds, col = espave, fill = espave)) + 
  geom_boxplot(alpha = 0.7, width = 0.5, size = 0.75) + xlab(" ") + ylab("Predicted Abundance") + coord_cartesian(ylim = c(0,5)) +
  scale_x_discrete(labels = c("Non-Espavé","Espavé")) +
  scale_color_manual("Tree Species", values = c("orange","grey20"), labels = c("Non-Espavé","Espavé")) + 
  scale_fill_manual("Tree Species", values = c("orange","grey78"), labels = c("Non-Espavé","Espavé")) + 
  theme_classic() + theme(legend.position = "none", axis.text.x = element_text(size = 10.92, color = "black", vjust = -4)); gg_count_espave

gg_count_dbh <- ggplot(global_dat_count_dbh, aes(x = dbh, y = preds, col = "All")) + 
  stat_smooth(size = 1, alpha = 0.5, method = "glm", method.args = list(family = "quasipoisson")) +
  xlab("Diameter at Breast Height (scaled)") + ylab("Predicted Abundance") + coord_cartesian(ylim = c(0,5)) +
  scale_color_manual("Tree Species", values = c("firebrick3"), labels = c("All")) + 
  theme_classic() + theme(legend.position = "none"); gg_count_dbh

count_3_pan <- ggarrange(gg_count_canopy_espave, gg_count_espave, gg_count_dbh,
          ncol = 1, labels = c("(d)","(e)","(f)"),
          label.x = 0.08, label.y = 0.985); 
```

# Merge predictions from presence and count models 
```{r}
ggarrange(pres_3_pan, count_3_pan, ncol = 2)
```



